### 2023-03-09 ESS_Olszewski/Olszewski.tex
Additional corrections added

\begin{artengenv}{Adam Olszewski}
	{Will a~human always outsmart a~computer? An essay}
	{Will a~human always outsmart a~computer? An essay}
	{Will a~human always outsmart a~computer? An essay}
	{Pontifical University of John Paul II in Kraków}
	{The title question of the paper has its empirical origin in the form of an individual's existential experience arising from the personal use of a~computer, which we attempt to describe in the first section. The rest of the entire paper can be understood as a~philosophical essay answering the question posed. First the connection between the main problem of the article and its ``premonition'' by mankind, which was expressed in the form of ancient myths and legends, is briefly suggested. After shortly discussing the problems that early considerations of AI focused on, i.e. whether machines can think at all, we move on to reformulate our title question, about the possibility of outsmarting AI. This outsmarting will be understood by us in a~rather limited way as to prevent a~machine from completing its implemented task. To achieve this objective, after softly clarifying the basic terms, an analogy is built between the ``outsmarting'' of a~machine by a~human (the target domain) and the playing of a~mathematical game between two players (the base domain), where this outsmarting is assigned a~``winning strategy'' in the certain game. This mathematical model is formed by games similar to Banach-Mazur games. The strict theorems of such games are then proved and applied to the target of the analogy. We then draw conclusions and look for counter-examples to our findings. The answer to the title question posed is negative, and it is not clear how far it should be taken seriously.
	}
	{existential experience, myth, computer, machine, Banach-Mazur games, winning strategy.}

@@

The meager and comeager sets relate to Baire space $\Box^\Box$. Intuitively ``comeager sets are large. They form a~filter, are dense, uncountable, and are closed under countable intersections. Meager sets are small. They form an ideal, and countable sets are meager.''
%\label{ref:RNDIAkiz5FrEH}(Soare, 2016, p.212).
\parencite[][p.212]{soare_turing_2016}. %
 I~mention this because the matter may be of interest to philosophers.
\begin{proposition}
Let $A=GR_{1}$, then there is a~winning strategy for Player \textbf{II}.
\end{proposition}
%\textbf{Corollary 7.4}. Let A=GR\textsubscript{1}, then there is a~winning strategy for Player \textbf{II}.
\begin{proof}
Let $A=GR_{1}$, i.e. $A$~is the set of all unary generally recursive functions. The set $GR_{1}$ is countable. Therefore, by virtue of theorem~\ref{olsz71}, there exists a~winning strategy for Player \textbf{II}.
\end{proof}
%Proof:
%
%Let A=GR\textsubscript{1}, i.e. A~is the set of all unary generally recursive functions. The set GR\textsubscript{1} is countable. Therefore, by virtue of theorem 7.1., there exists a~winning strategy for Player \textbf{II}.[F07F?]

\begin{proposition}
Let $A \subset \omega^{\omega}$ and $A = \{f: f \in GR_{1}$ and $f(n)=s$, for even $n$\}. Then Player \textbf{II} has a~winning strategy.
\end{proposition}
%\textbf{Corollary 7.5}. Let A~${\subset}$ ${\omega}$\textsuperscript{${\omega}$} and A~= \{f: f${\in}$GR\textsubscript{1} and f(n)=s, for even n\}. Then Player \textbf{II} has a~winning strategy.
\begin{proof}
\textls[-1]{$|A| < |\omega^{\omega}|$, then by virtue of theorem~\ref{olsz71} we have the thesis.}
\end{proof}
%Proof:
%
%{\textbar}A{\textbar} {\textless} {\textbar}${\omega}$\textsuperscript{${\omega}$}{\textbar}, then by virtue of theorem 7.1., we have the thesis.[F07F?]

\begin{proposition}
Let $A \subset \omega^{\omega}$ and $A = \{f: f(2n) = s\}$. Then Player~\textbf{I} has a~winning strategy.
\end{proposition}
%\textbf{Corollary 7.6}. Let A~${\subset}$ ${\omega}$\textsuperscript{${\omega}$} and A~= \{f: f(2n) = s\}. Then Player \textbf{I} has a~winning strategy.
\begin{proof}
Since $|A| = |\omega^{\omega}|$, we cannot use Theorem\ref{olsz71}. The winning strategy for Player~\textbf{I} consists in continuously choosing a~constant $s$, i.e, $\sigma (\langle x_{0}, y_{0}, x_{1}, y_{1}, \ldots, y_{i} \rangle) := s$, for any $i$. Let any $y\in\omega^{\omega}$ be a~sequence of the moves of Player~\textbf{II}. Then for each $y$, $\sigma^*y = z$, where $z := \langle x_{0}(=s), y_{0}, x_{1}(=s), y_{1}, \ldots \rangle$. Sequence $z$~has such a~form that for each $i$, $z(2i) = s = x_{i} = f(2i)$ for any function $f\in A$.
\end{proof}
%Proof:
%
%Since {\textbar}A{\textbar} = {\textbar}${\omega}$\textsuperscript{${\omega}$}{\textbar}, we cannot use Theorem 7.1. The winning strategy for Player \textbf{I} consists in continuously choosing a~constant s, i.e, ${\sigma}$(${\langle}$x\textsubscript{0}, y\textsubscript{0},x\textsubscript{1}, y\textsubscript{1}, \ldots, y\textsubscript{i}${\rangle}$) := s, for any i. Let any y${\in}{\omega}$\textsuperscript{${\omega}$} be a~sequence of the moves of Player \textbf{II}. Then for each y, ${\sigma}*$y = z, where z~:= ${\langle}$x\textsubscript{0}(=s), y\textsubscript{0}, x\textsubscript{1}(=s), y\textsubscript{1}, \ldots,${\rangle}$. Sequence z~has such a~form that for each i, z(2i) = s~= x\textsubscript{i} = f(2i) for any function f${\in}$A.[F08C?]

@@

Let us now formulate Theorem \ref{olszTY}, which---for obvious reasons---will not be precise in the case considered in the paper:
\begin{customthm}{T(Y)}\label{olszTY}
Any program (algorithm) implemented on a~machine will accomplish the goal written into the program, and a~\textit{ordinary user} cannot change it, which means that the user will always lose.
\end{customthm}
%\begin{enumerate}
%\item \begin{itemize}
%\item \textbf{Theorem T(Y}) Any program (algorithm) implemented on a~machine will accomplish the goal written into the program, and a~\textit{ordinary user} cannot change it, which means that the user will always lose.
%\end{itemize}
%\end{enumerate}
\section{Crackers}
Unlike Theorem T(X), which is a~mathematical theorem that is always true, Theorem T(Y) is empirical, and thus a~counter-example can be found for it. A~counter-example in such situations can be generated because of an unforeseen gap in the understanding of basic terms. A~group of unusual computer users, called \textit{crackers}, generate counter-examples to Theorem T(Y). The unusual nature of these users of a~machine, somewhat akin to hackers, is expressed in their setting a~goal for themselves to overcome the limitation implied by Theorem T(Y).\footnote{Note that a~Banach-Mazur game does not allow players to break the rules of the game.} There are essentially two ways in which crackers operate: breaking into a~program and breaking into a~server. Crackers are not ordinary computer users, they are often very knowledgeable and competent in certain areas of computer science, and their inexperienced followers are called \textit{script kiddies}. Crackers break firewalls, i.e. these features of a~program which are to ensure victory in the game to the algorithm, thus, essentially they break the rules of the game. Banach-Mazur games do not provide for such cases, although from the mathematical side this does not change anything because only the assumptions of a~theorem become unfulfilled, and the theorem becomes empty satisfied. Thus, in the context of the main question of the paper, we can make this optimistic prediction:
\begin{itemize}
\item For any algorithm, if an algorithm follows a~program written by a~human, then its \textit{cracker} exists.
\end{itemize}
A~pessimistic prediction referring to the notion of \textit{singularity} would sound like this:
\begin{itemize}
\item There exists such an algorithm, not necessarily written by a~human, that its \textit{cracker} does not exist.
\end{itemize}

>>>>

\begin{artengenv}{Adam Olszewski}
	{Will a~human always outsmart a~computer? An essay}
	{Will a~human always outsmart a~computer? An essay}
	{Will a~human always outsmart a~computer? An essay}
	{Pontifical University of John Paul II in Kraków, Poland}
	{The title question of the paper has its empirical origin in the form of an individual's existential experience arising from the personal use of a~computer, which we attempt to describe in the first section. The rest of the entire paper can be understood as a~philosophical essay answering the question posed. First the connection between the main problem of the article and its ``premonition'' by mankind, which was expressed in the form of ancient myths and legends, is briefly suggested. After shortly discussing the problems that early considerations of AI focused on, i.e. whether machines can think at all, we move on to reformulate our title question, about the possibility of outsmarting AI. This outsmarting will be understood by us in a~rather limited way as to prevent a~machine from completing its implemented task. To achieve this objective, after softly clarifying the basic terms, an analogy is built between the ``outsmarting'' of a~machine by a~human (the target domain) and the playing of a~mathematical game between two players (the base domain), where this outsmarting is assigned a~``winning strategy'' in the certain game. This mathematical model is formed by games similar to Banach-Mazur games. The strict theorems of such games are then proved and applied to the target of the analogy. We then draw conclusions and look for counter-examples to our findings. The answer to the title question posed is negative, and it is not clear how far it should be taken seriously.
	}
	{existential experience, myth, computer, machine, Banach-Mazur games, winning strategy.}

@@

The meager and comeager sets relate to Baire space $\omega^{\omega}$. Intuitively ``comeager sets are large. They form a~filter, are dense, uncountable, and are closed under countable intersections. Meager sets are small. They form an ideal, and countable sets are meager.''
%\label{ref:RNDIAkiz5FrEH}(Soare, 2016, p.212).
\parencite[][p.212]{soare_turing_2016}. %
 I~mention this because the matter may be of interest to philosophers.
\begin{proposition}
Let $A=GR_{1}$, then there is a~winning strategy for Player \textbf{II}.
\end{proposition}
%\textbf{Corollary 7.4}. Let A=GR\textsubscript{1}, then there is a~winning strategy for Player \textbf{II}.
\begin{proof}
Let $A=GR_{1}$, i.e. $A$~is the set of all unary general recursive functions. The set $GR_{1}$ is countable. Therefore, by virtue of theorem~\ref{olsz71}, there exists a~winning strategy for Player \textbf{II}.
\end{proof}
%Proof:
%
%Let A=GR\textsubscript{1}, i.e. A~is the set of all unary generally recursive functions. The set GR\textsubscript{1} is countable. Therefore, by virtue of theorem 7.1., there exists a~winning strategy for Player \textbf{II}.[F07F?]

\begin{proposition}
Let $A \subset \omega^{\omega}$ and $A = \{f: f \in GR_{1}$ and $f(n)=s$, for even $n\}$. Then Player \textbf{II} has a~winning strategy.
\end{proposition}
%\textbf{Corollary 7.5}. Let A~${\subset}$ ${\omega}$\textsuperscript{${\omega}$} and A~= \{f: f${\in}$GR\textsubscript{1} and f(n)=s, for even n\}. Then Player \textbf{II} has a~winning strategy.
\begin{proof}
\textls[-1]{$|A| < |\omega^{\omega}|$, then by virtue of theorem~\ref{olsz71} we have the thesis.}
\end{proof}
%Proof:
%
%{\textbar}A{\textbar} {\textless} {\textbar}${\omega}$\textsuperscript{${\omega}$}{\textbar}, then by virtue of theorem 7.1., we have the thesis.[F07F?]

\begin{proposition}
Let $A \subset \omega^{\omega}$ and $A = \{f: f(2n) = s\}$. Then Player~\textbf{I} has a~winning strategy.
\end{proposition}
%\textbf{Corollary 7.6}. Let A~${\subset}$ ${\omega}$\textsuperscript{${\omega}$} and A~= \{f: f(2n) = s\}. Then Player \textbf{I} has a~winning strategy.
\begin{proof}
Since $|A| = |\omega^{\omega}|$, we cannot use Theorem \ref{olsz71}. The winning strategy for Player~\textbf{I} consists in continuously choosing a~constant $s$, i.e, $\sigma (\langle x_{0}, y_{0}, x_{1}, y_{1}, \ldots, y_{i} \rangle) := s$, for any $i$. Let any $y\in\omega^{\omega}$ be a~sequence of the moves of Player~\textbf{II}. Then for each $y$, $\sigma^*y = z$, where $z := \langle x_{0}(=s), y_{0}, x_{1}(=s), y_{1}, \ldots \rangle$. Sequence $z$~has such a~form that for each $i$, $z(2i) = s = x_{i} = f(2i)$ for any function $f\in A$.
\end{proof}
%Proof:
%
%Since {\textbar}A{\textbar} = {\textbar}${\omega}$\textsuperscript{${\omega}$}{\textbar}, we cannot use Theorem 7.1. The winning strategy for Player \textbf{I} consists in continuously choosing a~constant s, i.e, ${\sigma}$(${\langle}$x\textsubscript{0}, y\textsubscript{0},x\textsubscript{1}, y\textsubscript{1}, \ldots, y\textsubscript{i}${\rangle}$) := s, for any i. Let any y${\in}{\omega}$\textsuperscript{${\omega}$} be a~sequence of the moves of Player \textbf{II}. Then for each y, ${\sigma}*$y = z, where z~:= ${\langle}$x\textsubscript{0}(=s), y\textsubscript{0}, x\textsubscript{1}(=s), y\textsubscript{1}, \ldots,${\rangle}$. Sequence z~has such a~form that for each i, z(2i) = s~= x\textsubscript{i} = f(2i) for any function f${\in}$A.[F08C?]

@@

Let us now formulate Theorem \ref{olszTY}, which---for obvious reasons---will not be precise in the case considered in the paper:
\begin{customthm}{T(Y)}\label{olszTY}
Any program (algorithm) implemented on a~machine will accomplish the goal written into the program, and an \textit{ordinary user} cannot change it, which means that the user will always lose.
\end{customthm}
%\begin{enumerate}
%\item \begin{itemize}
%\item \textbf{Theorem T(Y}) Any program (algorithm) implemented on a~machine will accomplish the goal written into the program, and a~\textit{ordinary user} cannot change it, which means that the user will always lose.
%\end{itemize}
%\end{enumerate}
\section{Crackers}
Unlike Theorem T(X), which is a~mathematical theorem that is always true, Theorem T(Y) is empirical, and thus a~counter-example can be found for it. A~counter-example in such situations can be generated because of an unforeseen gap in the understanding of basic terms. A~group of unusual computer users, called \textit{crackers}, generate counter-examples to Theorem T(Y). The unusual nature of these users of a~machine, somewhat akin to hackers, is expressed in their setting a~goal for themselves to overcome the limitation implied by Theorem T(Y).\footnote{Note that a~Banach-Mazur game does not allow players to break the rules of the game.} There are essentially two ways in which crackers operate: breaking into a~program and breaking into a~server. Crackers are not ordinary computer users, they are often very knowledgeable and competent in certain areas of computer science, and their inexperienced followers are called \textit{script kiddies}. Crackers break firewalls, i.e. these features of a~program which are to ensure victory in the game to the algorithm, thus, essentially they break the rules of the game. Banach-Mazur games do not provide for such cases, although from the mathematical side this does not change anything because only the assumptions of a~theorem become unfulfilled, and the theorem becomes empty satisfied. Thus, in the context of the main question of the paper, we can make this optimistic prediction:
\begin{itemize}
\item For any algorithm, if an algorithm follows a~program written by a~human, then its \textit{cracker} exists.
\end{itemize}
A~pessimistic prediction referring to the notion of \textit{singularity} would sound like this:
\begin{itemize}
\item There exists such an algorithm, not necessarily written by a~human, that its \textit{cracker} does not exist.
\end{itemize}

### 2023-03-09 ESS_Trzesicki/Trzesicki.tex
Additional corrections added

\begin{artengenv}{Kazimierz Trzęsicki}
	{Perspective on Turing paradigm:\\An essay}
	{Perspective on Turing paradigm: An essay}
	{Perspective on Turing paradigm: An essay}
	{University of Białystok}
	{Scientific knowledge is acquired according to some paradigm. Galileo wrote that the ``book of nature'' was written in mathematical language and could not be understood unless one first understood the language and recognized the characters with which it was written. It is argued that Turing planted the seeds of a new paradigm. According to the Turing Paradigm, the ``book of nature'' is written in algorithmic language, and science aims to learn how the algorithms change the physical, social, and human universe. Some sources of the Turing Paradigm are pointed out, and a few examples of the application of the Turing Paradigm are discussed.
	}
	{Galileo Galilei, Alan Turing, Konrad Zuse, zero, Arabic numeral, paradigm, mathematics, algorithmics.}

@@

The use of binary code was not obvious. Completed in the summer of 1946 American $ENIAC$, unlike binary coded $Z3$, $ABC$ and $Colossus$, was based on decimal arithmetic.

@@

Let us add that the word ``computer'' was still in the 19\textsuperscript{th} century, and even in 1936---when Turing published \emph{On Computable Numbers, with an Application to the {E}ntscheidungsproblem} \parencite*{Turing1936}---used to indicate an official who was doing cumbersome numerical calculations \parencite[p.446]{CopelandSprevakShagrir2017}. Thus understood, ``computer'' would denote a reckoner, in Polish: rachmistrz. Polish texts in which ``computer'' is translated into ``komputer'' are devoid of any associations present in English texts. In particular, the association of ``komputer'' with ``rachmistrz'' is important for the correct understanding of the  Turing texts.
\enlargethispage{1.5\baselineskip}

@@

 Zuse in \emph{Rechnender Raum} \parencite*{Zuse1967} is the first to talk about the universe as a computer network. He does not announce that he has a complete theory of everything in the form of some algorithm for counting the universe, but in this text he is the first to clearly formulate such an idea. He published the results of his further reflections in \emph{Rechnender Raum} \parencite*{Zuse1969} translated as \emph{Nature as Computation} \parencite* {Zuse2012}. In \emph {Der Computer} \parencite* {Zuse2010} he mentions: 
 \myquote{
%%PP Es geschah bei dem Gedanken der Kausalit{\"a}t, dass mir pl{\"o}tzlich der Gedanke auftauchte, den Kosmos als eine gigantische Rechenmaschine aufzufassen. Ich dachte dabei an die Relaisrechner: Relaisrechner enthalten Relaisketten. St{\"o}{\ss}t man ein Relais an, so pflanzt sich dieser Impuls durch die ganze Kette fort. So m{\"u}{\ss}te sich auch ein Lichtquant fortpflanzen, ging es mir durch den Kopf. Der Gedanke setzte sich fest; ich habe ihn im Laufe der Jahre zur Idee des ``Rechnenden Raumes'' ausgebaut. Es sollte freilich drei{\ss}ig Jahre dauern, ehe mir eine erste konkrete Formulierung der Idee gelang.    
Es geschah bei den Betrachtungen {\"u}ber die Kausalit{\"a}t, da{\ss} mir pl{\"o}tzlich der Gedanke auftauchte, den Kosmos als eine gigantische Rechenmaschine aufzufassen. Ich dachte dabei an die Relaisrechner: Relaisrechner enthalten Relaisketten. St{\"o}{\ss}t man ein Relais an, so pflanzt sich dieser Impuls durch die ganze Kette fort. So m{\"u}{\ss}te sich auch ein Lichtquant fortpflanzen, ging es mir durch den Kopf. Der Gedanke setzte sich fest; ich habe ihn im Laufe der Jahre zur Idee des ``Rechnenden Raumes'' ausgebaut. Es sollte freilich drei{\ss}ig Jahre dauern, ehe mir eine erste konkrete Formulierung der Idee gelang.
} 
 It was only in the third millennium that the idea of the world as a computer began to attract more attention. Among others, in \emph{Scientific American} and \emph{Spectrum der Wissenschaft} texts such as ``Is the universe a Big Computer?'', ``Is the Universe a Computer?'' are published. In the Autumn of 2006 the Technische Universit{\"a}t Berlin, where Horst Zuse, the son of Konrad Zuse then was a professor, organized a conference \emph{Ist das Universum ein Computer?} (Is the Universe a Computer?) \parencite[p.61]{German2012}.

>>>>

\begin{artengenv}{Kazimierz Trzęsicki}
	{Perspective on Turing paradigm:\\An essay}
	{Perspective on Turing paradigm: An essay}
	{Perspective on Turing paradigm: An essay}
	{University of Białystok, Poland}
	{Scientific knowledge is acquired according to some paradigm. Galileo wrote that the ``book of nature'' was written in mathematical language and could not be understood unless one first understood the language and recognized the characters with which it was written. It is argued that Turing planted the seeds of a new paradigm. According to the Turing Paradigm, the ``book of nature'' is written in algorithmic language, and science aims to learn how the algorithms change the physical, social, and human universe. Some sources of the Turing Paradigm are pointed out, and a few examples of the application of the Turing Paradigm are discussed.
	}
	{Galileo Galilei, Alan Turing, Konrad Zuse, zero, Arabic numeral, paradigm, mathematics, algorithmics.}

@@

The use of binary code was not obvious. Completed in the summer of 1946 American \textit{ENIAC}, unlike binary coded \textit{Z3}, \textit{ABC} and \textit{Colossus}, was based on decimal arithmetic.

@@

Let us add that the word ``computer'' was still in the 19\textsuperscript{th} century, and even in 1936---when Turing published \emph{On Computable Numbers, with an Application to the {E}ntscheidungsproblem} \parencite*{Turing1936}---used to indicate an official who was doing cumbersome numerical calculations \parencite[p.446]{CopelandSprevakShagrir2017}. Thus understood, ``computer'' would denote a reckoner, in Polish: ``rachmistrz''. Polish texts in which ``computer'' is translated into ``komputer'' are devoid of any associations present in English texts. In particular, the association of ``komputer'' with ``rachmistrz'' is important for the correct understanding of the  Turing texts.
\enlargethispage{1.5\baselineskip}

@@

 Zuse in \emph{Rechnender Raum} \parencite*{Zuse1967} is the first to talk about the universe as a computer network. He does not announce that he has a complete theory of everything in the form of some algorithm for counting the universe, but in this text he is the first to clearly formulate such an idea. He published the results of his further reflections in \emph{Rechnender Raum} \parencite*{Zuse1969} translated as
%\emph{Nature as Computation}
\emph{Calculating Space}
  \parencite* {Zuse2012}. In \emph {Der Computer} \parencite* {Zuse2010} he mentions: 
 \myquote{
%%PP Es geschah bei dem Gedanken der Kausalit{\"a}t, dass mir pl{\"o}tzlich der Gedanke auftauchte, den Kosmos als eine gigantische Rechenmaschine aufzufassen. Ich dachte dabei an die Relaisrechner: Relaisrechner enthalten Relaisketten. St{\"o}{\ss}t man ein Relais an, so pflanzt sich dieser Impuls durch die ganze Kette fort. So m{\"u}{\ss}te sich auch ein Lichtquant fortpflanzen, ging es mir durch den Kopf. Der Gedanke setzte sich fest; ich habe ihn im Laufe der Jahre zur Idee des ``Rechnenden Raumes'' ausgebaut. Es sollte freilich drei{\ss}ig Jahre dauern, ehe mir eine erste konkrete Formulierung der Idee gelang.    
Es geschah bei den Betrachtungen {\"u}ber die Kausalit{\"a}t, da{\ss} mir pl{\"o}tzlich der Gedanke auftauchte, den Kosmos als eine gigantische Rechenmaschine aufzufassen. Ich dachte dabei an die Relaisrechner: Relaisrechner enthalten Relaisketten. St{\"o}{\ss}t man ein Relais an, so pflanzt sich dieser Impuls durch die ganze Kette fort. So m{\"u}{\ss}te sich auch ein Lichtquant fortpflanzen, ging es mir durch den Kopf. Der Gedanke setzte sich fest; ich habe ihn im Laufe der Jahre zur Idee des ``Rechnenden Raumes'' ausgebaut. Es sollte freilich drei{\ss}ig Jahre dauern, ehe mir eine erste konkrete Formulierung der Idee gelang.
} 
 It was only in the third millennium that the idea of the world as a computer began to attract more attention. Among others, in \emph{Scientific American} and \emph{Spectrum der Wissenschaft} texts such as ``Is the universe a Big Computer?'', ``Is the Universe a Computer?'' are published. In the Autumn of 2006 the Technische Universit{\"a}t Berlin, where Horst Zuse, the son of Konrad Zuse then was a professor, organized a conference \emph{Ist das Universum ein Computer?} (Is the Universe a Computer?) \parencite[p.61]{German2012}.

### 2023-03-09 REV_Mscislawski/Mscislawski.tex
Additional corrections added

\begin{newrevengenv}{Łukasz Mścisławski}
	{Is AI case that is explainable, intelligible or hopeless?}
	{Is AI case that is explainable, intelligible or hopeless?}
	{Is AI case that is explainable, intelligible or hopeless?}
	{Wrocław University of Science and Technology}
	{H. Cappelen, J. Dever, \textit{Making AI Intelligible: Philosophical Foundations}, Oxford University Press, Oxford 2021, 175+viii~pp.}
%	 183 pp. (175+viii).}

@@

\enlargethispage{.5\baselineskip}
In Chapter 4 (\textit{Our Theory. De-Anthropocentrized Externalism}), the authors make a~presentation of their own conception, which they describe as de-anthopocentrised externalism. They base their proposal on two basic claims: 1) the content of AI systems should be explained externalistically; and 2) existing externalist approaches are anthropocentric. The first thesis is based on the observation that content, related to action, is not a~problem at the level of software or computation. It is a~problem at the environmental and sociological level. The second thesis is the observation that however philosophers have developed impressive models of human language and human mental states. However, this is not the case with AI systems---the operation of software on specific hardware, both in the computational layer and in terms of the functioning of the hardware, is fundamentally different from what can be described by such means. A~de-anthropocentised metasemantics is therefore needed here (pp.59--71). However, some additional rule of thumb is also needed for the future selection of appropriate measures and the development of an effective content theory of AI systems. Here, the authors propose a~meta-metasemantic principle: interpreter-centric knowledge-maximisation. Two important issues also arise here, which can also serve as a~kind of guideline in the search for appropriate tools for further research: a) it is the human knowledge and not the AI system that is important, so the idea is to maximise human knowledge; and b) the perspective of interests is important here, bearing in mind that human interests may differ from those of the an AI system\footnote{In simple terms: a~human may want to know why he or she was classified negatively in given aspect, while an AI system may seek to optimise the data in some way (e.g. finding the minimum of a~function).} (pp.75--79).

The background thus outlined serves the authors to attempt to apply already existing philosophical tools when it comes to relating content to the results provided by AI systems. The test task is to classify a~particular person into a~particular category. The basic problematic therefore involves AK1) referring to that particular person; AK2) attributing to that person a~test characteristic (adjudicating that person as having that characteristic); AK3) criteria for classifying that person into a~particular category (adjudicating attribution to a~category). Chapter 5 (\textit{Application. The Predicate 'High Risk}') attempts to unravel these issues using an externalist approach based on proposals of Kripke. In doing so, they draw attention to the fundamental difficulties of such approaches: the problem of the anchoring event, the problem of defining chain of transmission and issuses connected with the problem of being part of communicative chain, when AI systems are involved (p.82-88). Additional difficulties are posed in relation to AK3 by the possible variability of classifications and models and the fact of context dependence. Unsurprisingly, it is very important to note that in the case of systems based on machine learning, the final correctness of the answers given depends on those that the human training such systems deems to be true, i.e. on human decisions. Another problem is that AI systems do not seem to have capacity to representing that could be analogous to human's ability of representing using proper names. Such a~situation generates serious problems of communicative and epistemic nature (pp.99--105). Cappalen and Dever make an analogous analysis when it comes to the potential application of the Mental Files Framework\footnote{Murez and Recanti shall be characterized as \textit{devices of direct reference whose deployment makes it possible to entertain singular thoughts, i.e. thoughts that are about particular objects rather than about whatever possesses certain features or satisfies such and such a~description}
%\label{ref:RNDzw4sL5zarJ}(cf. Murez and Recanati, 2016, p.267).
\parencite[cf.][p.267]{murez_mental_2016}.%
} and attempt to extend the findings of Evans and Recanati to Epistemically Rewarding Relations. Chapter 6 (\textit{Application. Names and the Mental Files Framework}) is devoted to this. The addition of the knowledge-maximisation rule to the framework in question raises the question of whether it is about maximising general or specific knowledge (p.114). Ultimately, however, it appears that with the philosophical tool in question the case is similar to that of the Kripke-style framework, a~simple application to AI systems may not be feasible. In particular, there is need to abstract the notion of an epistemically rewarding relationship. The main difficulties in the context of the philosophical framework in question, however, are the need to focus on particular epistemic goals and activities and the fact that what the results provided by an AI system are about depends on the aims of the interpreter. Hence the results of the considerations in Chapters 4 and 5 are reinforced and shows the organic nature of the internalist view of AI: you cannot bite off all the facts about the classification content of a~machine learning system by looking only at the internal implementation of that system (pp.115--116).

@@

The starting point for the strategy proposed by the authors, as could be seen in the brief presentation of the individual chapters, is an attempt to use attribution mechanisms that appear to be human-specific (which is the case in Chapter 5). However, the authors are aware that this kind of tactic cannot be applied across the board, as witnessed in Chapter 4, and the de-anthropocentric perspective proposed therein. However, when, as they rightly point out, this path does not yield satisfactory results, they change the tools with which they try to get their way (Chapter 6 and later Chapter 7). While this is understandable, it seems that the link between abandoning the previous path of trying to deal with the problem under investigation and the choice of subsequent tools is not sufficiently justified. A~side effect of such a~situation may be a~feeling that the reading of the book is piecewise smooth. The authors also do not make any remarks, whether the failure of the given tool in question is a~permanent or whether it is only temporary situation and we need more research in given area or wait for more advanced technologies\footnote{The answer to that question seems to be particularly in case of explainable AI.}.

>>>>

\begin{newrevengenv}{Łukasz Mścisławski}
	{Is AI case that is explainable, intelligible or hopeless?}
	{Is AI case that is explainable, intelligible or hopeless?}
	{Is AI case that is explainable, intelligible or hopeless?}
	{Wrocław University of Science and Technology, Poland}
	{H. Cappelen, J. Dever, \textit{Making AI Intelligible: Philosophical Foundations}, Oxford University Press, Oxford 2021, 175+viii~pp.}
%	 183 pp. (175+viii).}

@@

\enlargethispage{.5\baselineskip}
In Chapter 4 (\textit{Our Theory. De-Anthropocentrized Externalism}), the authors make a~presentation of their own conception, which they describe as de-anthopocentrised externalism. They base their proposal on two basic claims: 1) the content of AI systems should be explained externalistically; and 2) existing externalist approaches are anthropocentric. The first thesis is based on the observation that content, related to action, is not a~problem at the level of software or computation. It is a~problem at the environmental and sociological level. The second thesis is the observation that however philosophers have developed impressive models of human language and human mental states. However, this is not the case with AI systems---the operation of software on specific hardware, both in the computational layer and in terms of the functioning of the hardware, is fundamentally different from what can be described by such means. A~de-anthropocentised metasemantics is therefore needed here (pp.59--71). However, some additional rule of thumb is also needed for the future selection of appropriate measures and the development of an effective content theory of AI systems. Here, the authors propose a~meta-metasemantic principle: interpreter-centric knowledge-maximization. Two important issues also arise here, which can also serve as a~kind of guideline in the search for appropriate tools for further research: a) it is the human knowledge and not the AI system that is important, so the idea is to maximise human knowledge; and b) the perspective of interests is important here, bearing in mind that human interests may differ from those of the an AI system\footnote{In simple terms: a~human may want to know why he or she was classified negatively in given aspect, while an AI system may seek to optimise the data in some way (e.g. finding the minimum of a~function).} (pp.75--79).

The background thus outlined serves the authors to attempt to apply already existing philosophical tools when it comes to relating content to the results provided by AI systems. The test task is to classify a~particular person into a~particular category. The basic problematic therefore involves AK1) referring to that particular person; AK2) attributing to that person a~test characteristic (adjudicating that person as having that characteristic); AK3) criteria for classifying that person into a~particular category (adjudicating attribution to a~category). Chapter 5 (\textit{Application. The Predicate 'High Risk}') attempts to unravel these issues using an externalist approach based on proposals of Kripke. In doing so, they draw attention to the fundamental difficulties of such approaches: the problem of the anchoring event, the problem of defining chain of transmission and issuses connected with the problem of being part of communicative chain, when AI systems are involved (p.82-88). Additional difficulties are posed in relation to AK3 by the possible variability of classifications and models and the fact of context dependence. Unsurprisingly, it is very important to note that in the case of systems based on machine learning, the final correctness of the answers given depends on those that the human training such systems deems to be true, i.e. on human decisions. Another problem is that AI systems do not seem to have capacity to representing that could be analogous to human's ability of representing using proper names. Such a~situation generates serious problems of communicative and epistemic nature (pp.99--105). Cappalen and Dever make an analogous analysis when it comes to the potential application of the Mental Files Framework\footnote{Murez and Recanti shall be characterized as \textit{devices of direct reference whose deployment makes it possible to entertain singular thoughts, i.e. thoughts that are about particular objects rather than about whatever possesses certain features or satisfies such and such a~description}
%\label{ref:RNDzw4sL5zarJ}(cf. Murez and Recanati, 2016, p.267).
\parencite[cf.][p.267]{murez_mental_2016}.%
} and attempt to extend the findings of Evans and Recanati to Epistemically Rewarding Relations. Chapter 6 (\textit{Application. Names and the Mental Files Framework}) is devoted to this. The addition of the knowledge-maximization rule to the framework in question raises the question of whether it is about maximising general or specific knowledge (p.114). Ultimately, however, it appears that with the philosophical tool in question the case is similar to that of the Kripke-style framework, a~simple application to AI systems may not be feasible. In particular, there is need to abstract the notion of an epistemically rewarding relationship. The main difficulties in the context of the philosophical framework in question, however, are the need to focus on particular epistemic goals and activities and the fact that what the results provided by an AI system are about depends on the aims of the interpreter. Hence the results of the considerations in Chapters 4 and 5 are reinforced and shows the organic nature of the internalist view of AI: you cannot bite off all the facts about the classification content of a~machine learning system by looking only at the internal implementation of that system (pp.115--116).

@@

The starting point for the strategy proposed by the authors, as could be seen in the brief presentation of the individual chapters, is an attempt to use attribution mechanisms that appear to be human-specific (which is the case in Chapter 5). However, the authors are aware that this kind of tactic cannot be applied across the board, as witnessed in Chapter 4, and the de-anthropocentric perspective proposed therein. However, when, as they rightly point out, this path does not yield satisfactory results, they change the tools with which they try to get their way (Chapter 6 and later Chapter 7). While this is understandable, it seems that the link between abandoning the previous path of trying to deal with the problem under investigation and the choice of subsequent tools is not sufficiently justified. A~side effect of such a~situation may be a~feeling that the reading of the book is piecewise smooth. The authors also do not make any remarks, whether the failure of the given tool in question is a~permanent or whether it is only temporary situation and we need more research in given area or wait for more advanced technologies.\footnote{The answer to that question seems to be particularly in case of explainable AI.}
